---
title: "Bay Area Accidents"
date: "12/2/2024"
author: "Sean Steven Alcantara"
format:
  html:
    embed-resources: true
output: 
  flexdashboard::flex_dashboard:
---

```{r}
library(tidyverse)
library(mapproj)
library(maps)
library(leaflet)
```

```{r}
road_accidents <- read.csv("../Project/US_Accidents_March23.csv")
```

```{r}
accidents <- road_accidents %>% 
  select(Severity, Start_Time, State, County, Start_Lat, Start_Lng,
         Weather_Condition, Amenity:Turning_Loop) %>% 
  mutate(Start_Time = ymd_hms(Start_Time)) %>% 
  mutate(Severity = as.character(Severity),
         year = year(Start_Time),
         date = date(Start_Time),
         month = month(Start_Time, label = TRUE),
         wday = wday(Start_Time, label = TRUE),
         hour = hour(Start_Time)) %>% 
  filter(! year %in% c(2016, 2023))
```

```{r}
accidents[,8:20] <- apply(accidents[,8:20], MARGIN = 2, FUN = as.logical)
```

```{r}
bay_area <- c("Alameda", "Contra Costa", "Sonoma", "Napa", "Solano",
              "Marin", "San Francisco", "San Mateo", "Santa Clara")
accidents_bay_area <- accidents %>% 
  filter(State == "CA", County %in% bay_area)
```

```{r}
accidents_bay_area_2020_Jul_Oct<- accidents_bay_area %>% 
  filter(year %in% c(2020),
         month %in% c("Jul", "Aug", "Sep", "Oct"))
```

```{r}
accidents_bay_area_2020and2021_Jul_Oct<- accidents_bay_area %>% 
  filter(year %in% c(2020, 2021),
         month %in% c("Jul", "Aug", "Sep", "Oct"))
```

```{r}
accident_count <- nrow(accidents_bay_area)
POI_accidents <- accidents_bay_area %>%
  select(Amenity:Turning_Loop, Severity) %>% 
  pivot_longer(Amenity:Turning_Loop, names_to = "POI") %>% 
  select(POI, value, Severity) %>% 
  mutate(POI = as.factor(POI))
```

```{r}
bay_area_lowercase <- tolower(bay_area)
bay_area_map <- map_data("county", "california") %>% 
  filter(subregion %in% bay_area_lowercase)
```

```{r setup, include=FALSE}
library(flexdashboard)
```

# Part 1

Row
--------------------------------------------------------------------
**(a) Question: What is the source of your data?**
 Source:
 <https://www.kaggle.com/datasets/sobhanmoosavi/us-accidents?resource=download>
 Description:
 This is a countrywide car accident dataset that covers 49 states of the USA. The accident data were collected from February 2016 to March 2023,
 using multiple APIs that provide streaming traffic incident (or event) data. These APIs broadcast traffic data captured by various entities,
 including the US and state departments of transportation, law enforcement agencies, traffic cameras, and traffic sensors within the road networks.
 The dataset currently contains approximately 7.7 million accident records.


Row
--------------------------------------------------------------------
**(b) Question: For each variable, state what type.**
 
-   categorical

    -   Severity: hows the severity of the accident, a number between 1 and 4, where 1 indicates the least impact on traffic (i.e., short delay as a result of the accident) and 4 indicates a significant impact on traffic (i.e., long delay).
    -   Weather_Condition: Shows the weather condition (rain, snow, thunderstorm, fog, etc.)
    -   POI: Points of interest near an accident. i.e. Junction, Traffic_Light, Crossing, etc.
    -   State: State where the accident happened
    -   County: County where the accident happened

-   time

    -   Start_Time: Time when the accident happened

-   quantitative

    -   Distance.mi.: The length of the road extent affected by the accident in miles.

-   geolocation

    -   Start_Lat: Shows latitude in GPS coordinate of the start point.
    -   Start_Lng: Shows longitude in GPS coordinate of the start point.

# Part 2

## (a)

**(a) Question: Make list of the visualization you have created for your dashboard. State whether you have reproduced or modified some existing visualization, provide the links to the original visualization.**

I am making my own visualizations.

-   3 Barplot

-   1 Map

# Plots {.storyboard}

### Bay Area Accidents Per Year

```{r}
ggplot(accidents_bay_area, aes(x = month)) +
  geom_bar(col = "black") +
  facet_wrap(vars(year), ncol = 2) +
  guides(x = guide_axis(angle = 90)) +
  ggtitle("Bay Area Accidents")
```

***
We can see that there's an increase in the count of accidents after 2020, compared to before 2021. We also see
a large dip in accidents around July to October of 2020. This is likely an effect of the Covid lockdowns that started on April 2020

### Bay Area Hourly Accidents July to Oct (2020 vs 2021)

```{r}
ggplot(accidents_bay_area_2020and2021_Jul_Oct, aes(x = hour)) +
  scale_fill_brewer(palette = "Blues") +
  geom_histogram(aes(group = Severity, fill = Severity),
                 col = "black", bins = 24) +
  facet_wrap(vars(year), ncol = 2) +
  ggtitle("Bay Area hourly accidents July to Oct (2020 vs 2021)")
```

***

### Bay Area Accidents near Points of Interest

```{r}
POI_accidents %>%
  filter(value == TRUE) %>% 
  ggplot(aes(y = fct_rev(fct_infreq(POI)))) +
  scale_fill_brewer(palette = "BuGn") +
  geom_bar(aes(fill = Severity), position = "stack", col = "black") +
  ylab("POI") +
  ggtitle("Bay Area Accidents near a Point of Interest (POI)")
```
# Map of Bay Area Accidents

### Map of Bay Area Accidents (July-Oct 2020)

```{r}
leaflet(data = accidents_bay_area_2020_Jul_Oct) %>% 
  addTiles() %>% 
  addCircleMarkers(
    lng = ~Start_Lng,
    lat = ~Start_Lat,
    popup = ~paste(Start_Time, "<br>", "Severity =", Severity),
    radius = 1
  )
```

# Part 4

## (a)

**(a) Question: Write a narrative explaining the story of your data, visualizations and your dashboard.**
