---
title: "STAT450, Homework 3"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

# Exercise 1

```{r}
library(tidyverse)
```

# (a)

```{r}
ggplot(data = mpg, mapping = aes(y = cty, x = displ, color = drv)) +
  geom_point() + geom_smooth(method="lm", se = FALSE)
```

# (b)

```{r}
ggplot(data = mpg, mapping = aes(y = cty, x = displ)) +
  geom_point() + geom_smooth(method="lm", se = FALSE) +
  facet_wrap(facets = vars(drv))
```

# (c)

```{r}
ggplot(data = mpg, mapping = aes(y = cty, x = displ)) +
  geom_point(aes(color = drv)) + geom_smooth(method = "lm", se = FALSE, col = "black")
```

# (d)

```{r}

ggplot(data = mpg, mapping = aes(y = manufacturer, fill = class)) +
  geom_bar()
```

# Exercise 2

```{r}
library(mosaicData)
```

```{r}
head(CPS85)
```

```{r}
dim(CPS85)
help("CPS85")
```

# (a)

```{r}
ggplot(data = CPS85, mapping = aes(x = wage)) +
  geom_histogram(binwidth = 3, col = "white", fill = "blue") 
```

```{r}
ggplot(data = CPS85) +
  geom_density(mapping = aes(x = wage), col = "blue")
```

The histogram and density plots for wage is unimodal, and it is skewed right

# (b)

```{r}
ggplot(data = CPS85, mapping = aes(y = wage, x = sector)) +
  geom_boxplot()
```

The two sectors with the highest median wage are "prof" and "manuf".\
The "manuf" sector has the greatest variability in wage

# (c)

```{r}
ggplot(data = CPS85, mapping = aes(x = sector)) +
  geom_bar()
```

The "prof" sector has the highest number of employees. There are approximately 110 employees in the "prof" sector.

# (d)

```{r}
ggplot(data = CPS85, mapping = aes(x = sector, fill = sex)) +
  geom_bar()
```

# (e)

```{r}
ggplot(data = CPS85, mapping = aes(x = sector, fill = sex)) +
  geom_bar(position = "fill")
```

The "prof" sector have roughly the same proportion of female and male employees. The sales sector is the next sector with the closest proportion of female and male employees.

# Exercise 3

```{r}
library(maps)
library(mapproj)

```

# (a)

```{r}
ca <- map_data("county", "ca")
ggplot(data = ca, aes(x = long, y = lat, group = group)) +
geom_polygon(fill = "white", color = "black") + coord_map()
```

# (b)

```{r}
head(ca)
```

```{r}
unique(ca$subregion)
length(unique(ca$subregion))
```

Explanation:

unique(ca\$subregion) –\> creates a vector filled with unique values from the ca\$subregion.\
length(unique(ca\$subregion)) –\> how many unique values there are in ca\$subregion

# (c)

```{r}
alameda_ca <- filter(ca, subregion == "alameda")
as_tibble(alameda_ca)
```

# (d)

```{r}
ggplot(data = alameda_ca, mapping = aes(x = long, y = lat, group = group)) +
  geom_polygon(fill = "white", color = "black") + coord_map()
```

# Bonus

```{r}
bay_area <- filter(ca, subregion == "alameda" | subregion == "contra costa" |
                       subregion == "marin" | subregion == "napa" |
                       subregion == "san francisco" | subregion == "san mateo" |
                       subregion == "santa clara" | subregion == "solano" |
                       subregion == "sonoma")
unique(bay_area$subregion)
length(unique(bay_area$subregion))
```

```{r}
ggplot(data = bay_area, mapping = aes(x = long, y = lat, group = group, )) +
  geom_polygon(mapping = aes(fill = subregion), color = "black") +
  coord_map()
```
