---
title: "STAT450, Homework 2"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
cdc <- read.csv("https://ericwfox.github.io/data/cdc1000.csv")
head(cdc)
```

# Exercise 1

```{r}
exerany_table <- table(cdc$exerany)
barplot(exerany_table, names.arg = c("no", "yes"),
        xlab = "Exercised in the last Month")
```

# Exercise 2

```{r}
plot(cdc$height ~ cdc$wtdesire, ylab = "Height (inches)",
     xlab = "Weight Desired (lbs)")
```

Explain association:

There is a positive association between Height and Weight Desired. It seems like there is also a linear association between the two

# Exercise 3

```{r}
genhlth_fac <- factor(x = cdc$genhlth,
                      levels = c("poor", "fair", "good", "very good", "excellent"))
boxplot(cdc$weight ~ genhlth_fac, ylab = "Weight", xlab = "General Health")
```

# Exercise 4

```{r}
wtdiff <- cdc$wtdesire - cdc$weight
summary(wtdiff)
```

```{r}
hist(wtdiff, xlab = "Difference(lbs)",
        main = "Difference in Weight Desired and Current Weight")
```

```{r}
boxplot(wtdiff, horizontal = TRUE, xlab = "Difference(lbs)",
        main = "Difference in Weight Desired and Current Weight")
```

Explain:

The dataset is unimodal since there is only one peak, and it is slightly left-skewed. There are also outliers since there are values that lie outside the whiskers of the box plot.
