---
title: "STAT450, Homework 4"
author: "Sean Steven C. Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
library(tidyverse)
library(nycflights13)
```

# Exercise 1

```{r}
hate_crimes <- read_csv("hate_crimes.csv")
head(hate_crimes)
```

## (a)

```{r}
# 10 states with the higest gini_index
hate_crimes |> 
  select(state, gini_index) |> 
  arrange(desc(gini_index)) |>
  head(n = 10)
```

```{r}
# 10 states with the lowest gini_index
hate_crimes |> 
  select(state, gini_index) |> 
  arrange(gini_index) |>
  head(n = 10)
```

## (b)

```{r}
ggplot(hate_crimes, aes(x = gini_index, y = avg_hatecrimes_per_100k_fbi)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE, method = "loess") +
  ylab(label = "Average annual hate crimes per 100,000 residents") +
  xlab("GINI Index")
```

There isn't a clear association between the GINI Index and the Average annual hate crime per 100,000 residents. There is a clear outlier and this is the District of Columbia with a GINI_Index value of 0.532.

# Exercise 2

## (a)

**Answer:**

read_csv2() is used to read data that use ";" to separate values

## (b)

**Answer:**

read_delim() is used to read data that uses "\|" to separate values. You just need to specify delim = "\|" as an argument to the function

# Exercise 3

```{r}
head(relig_income)
```

```{r}
names(relig_income)[-1]
```

```{r}
relig_income |> 
  pivot_longer(cols = names(relig_income)[-1],
               names_to = "income", values_to = "count")
```

# Exercise 4

## (a)

This is an inner join. Only the rows that are in both table were kept

## (b)

This is a left join. All the rows in table 1 (the left one) were kept

# Exercise 5

```{r}
head(planes)
```

```{r}
planes |> 
  count(tailnum) |> 
  filter(n > 1)
```

tailnum is a unique identifier for planes dataframe

# Exercise 6

```{r}
options(digits = 2)
```

```{r}
airports_location <- flights |> 
  group_by(dest) |> 
  summarize(count = n(), arr_delay_mean = mean(arr_delay, na.rm = TRUE)) |> 
  left_join(airports, by = c("dest" = "faa"))
airports_location
```

# Bonus

```{r}
library(maps)
library(mapproj)

```

```{r}
states <- map_data("state")
```

```{r}
airports_location |> 
  filter(!(dest %in% c("HNL", "ANC"))) |> 
  drop_na(lat) |> 
  ggplot(data = _, aes(x = lon, y = lat, col = arr_delay_mean)) +
  geom_polygon(data = states, aes(x = long, y = lat, group = group),
               fill = "white", color = "black") +
  geom_point(size = 2) +
  ggtitle("Average arrival delay per Destination Airport for nycflights13") +
  ylab("Latitude") +
  xlab("Longitude") +
coord_map()
```
