---
title: "STAT450, Exercise Lecture 12"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
library(tidyverse)
```

# Exercise 1

**Use the procedure described on the previous page to read the hate_crimes.csv data set into R.**

```{r}
hate_crimes <- read_csv("hate_crimes.csv")
```

# Exercise 2

**Compute some summary statistics for avg_hatecrimes_per_100k_fbi**

```{r}
head(hate_crimes)
```

```{r}
summary(hate_crimes)
```

# Exercise 3

**Which states had the highest hate crime rates?**

```{r}
hate_crimes |> 
  select(state, avg_hatecrimes_per_100k_fbi) |> 
  arrange(desc(avg_hatecrimes_per_100k_fbi)) |> 
  head(5)
```

**Which states had the lowest hate crime rates?**

```{r}
hate_crimes |> 
  select(state, avg_hatecrimes_per_100k_fbi) |> 
  arrange(avg_hatecrimes_per_100k_fbi) |> 
  head(5)
```

# Exercise 4

**Use ggplot() to make a scatter plot with share_unemployed_seasonal on the x-axis and avg_hatecrimes_per_100k_fbi on the y-axis. Use geom_smooth() to add a smooth trend line to scatter plot. Label the x-axis “Seasonally adjusted unemployment”, and the y-axis**

**“Average hate crimes per 100,000 residents, 2010-2015”. Describe any trends in the scatter plot.**

```{r}
ggplot(hate_crimes,
       aes(y = avg_hatecrimes_per_100k_fbi, x = share_unemployed_seasonal)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE) +
  xlab("Seasonally adjusted unemployment") + 
  ylab("Average hate crimes per 100,000 residents, 2010-2015")
```

There are no obvious trends. It seems like there is no relationship between the two variables.

```{r}
cor(x = hate_crimes$share_unemployed_seasonal,
    y = hate_crimes$avg_hatecrimes_per_100k_fbi,
    use = "complete.obs")
```

```{r}
ggplot(hate_crimes,
       aes(y = avg_hatecrimes_per_100k_fbi, x = gini_index)) +
  geom_point(size = 1) +
  geom_smooth(se = FALSE) +
  xlab("Seasonally adjusted unemployment") + 
  ylab("Average hate crimes per 100,000 residents, 2010-2015")
```
