---
title: "STAT450, Lecture 5 Exercise"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
cdc <- read.csv("https://ericwfox.github.io/data/cdc1000.csv")
cdc
```

```{r}
barplot(table(cdc$smoke100), names.arg = c("No", "Yes"),
        xlab = "Smoked at least 100 cigarettes", ylab = "Counts")
```

# Exercise 1

```{r}
addmargins(table(gender = cdc$gender, smoke100 = cdc$smoke100))
```

According to the contingency table above, what proportion of males have smoked at least 100 cigarettes

```{r}
252/488 * 100
```

51.6% of males have smoked at least 100

# Exercise 2

Create a bar plot of genhlth with categories ordered as poor, fair, good, very good, excellent

```{r}
gen_health <- factor(levels = c("poor", "fair", "good", "very good", "excellent"),
                     cdc$genhlth)
barplot(table(gen_health), ylab = "Count", xlab = "Health Level",)
```

# Exercise 3

Make a histogram and box plot for wtdesire. Describe the shape of the distribution

```{r}
boxplot(cdc$wtdesire, main = "Weight Desired",
        horizontal = TRUE, xlab = "weight desired(lbs)")
hist(cdc$wtdesire, main = "Weight Desired", xlab = "weight desired(lbs)", )
```

Right-skewed distribution

# Random

```{r}
gender <- factor(cdc$gender, levels = c("m, f"))
gender <- cdc$gender |> if ( . == "m" then)
plot(cdc$height ~ cdc$wtdesire, col = gender,
     xlab = "Height(inches)", ylab = "Weight Desired(lbs)")
```
