---
title: "Lecture 9, Exercises"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
library(tidyverse)
library(nycflights13)
```

```{r}
head(flights)
colnames(flights)
```

# Exercise 1

Use filter() to find all flights that • Had an arrival delay of two or more hours. • Were operated by United (UA), American (AA), or Delta (DL).

```{r}
flights |>
  select(arr_delay, dep_delay, carrier) |>
  mutate(tot_delay_hrs = (dep_delay + arr_delay) / 60) |>
  filter(tot_delay_hrs >= 2, carrier == "UA" | carrier == "AA" | carrier == "DL")
```

# Exercise 2

Run the following code. Which three variables get selected by contains("dep"), and how are they related?

```{r}
flights |>
select(year:day, carrier, contains("dep")) |> filter(carrier == "UA") |>
  arrange(desc(dep_delay))
```

Contains("dep") captured all columns that contain the substring "dep". In this case dep_time, sched_dep_time, and dep_delay.

# Exercise 3

How many flights have a missing value for dep_time? What other variables have missing values? What might these rows represent? (Hint: use is.na())

```{r}
flights |>
  filter(is.na(dep_time))
```

8255 NA values in flights\$dep_time. This means these flights have been cancelled

# Exercise from tibbles.pdf

Type the following into a code chunk in Quarto, and then render to HTML. How does printing a traditional data.frame in Quarto differ from printing a tibble?

```{r}
iris # print iris data.frame
class(iris)
iris_tb <- as_tibble(iris) # convert iris to a tibble iris_tb # print iris tibble
iris_tb
class(iris_tb)
```

data.frame is printed in entirety in html

tibble prints something like head(n=10)
