---
title: "STAT450, Exercise, Lecture 10"
author: "Sean Steven Alcantara"
format: 
  html:
    self-contained: true
editor: visual
---

```{r}
library("tidyverse")
library("nycflights13")
head(mpg)
```

# Exercise 1

Using the mpg data frame, for each car manufacturer, compute the count, average city miles per gallon, and average highway miles per gallon.

```{r}
mpg |> group_by(manufacturer) |> 
  summarize(count = n(), mean_mpg_cty = mean(cty), mean_mpg_hwy = mean(hwy))
```

# Exercise 2

For each carrier, compute the number of flights and median departure delay. Which carrier has the longest median departure delay?

```{r}
flights |> group_by(carrier) |>
  summarize(count = n(), median_dep_delay = median(dep_delay, na.rm = TRUE))
```

FL is the carrier with the highest median departure delay.

# Exercise 3

For each carrier, compute the proportion of flights that were delayed more than 5 minutes (i.e., the departure delay was greater than 5 minutes). Which carrier has the smallest proportion of delays?

```{r}
flights |> group_by(carrier) |> 
  summarize(delay_proportion = mean(dep_delay > 5, na.rm = TRUE)) 
```

The HA carrier has the smallest proportion of delays
